\section{Advanced component generators}
\label{sec:impl2}
In this Chapter I introduce more advanced component generators for caching,
session management, load balancing, network communication and error handling.

\subsection{Storing data}
In order to implement behaviors such as caching or session management
components needs to be able to maintain an internal state.
To do this they need to store some information (this can be in memory, 
in a database etc.).
This section introduces an interface that specifies methods that an 
object must implement, in order to be used as a data storage. This 
separates the problem of managing the data storage from the logic of
the component that uses it.

\subsubsection{Storage Values}
Most of the components need to remove old values when they are no longer
valid or relevant. Hence, the package stores timestamped values. 
This values are represented by the struct type shown in Figure 
\ref{fig:StorageValue} below. There \texttt{Time} represents the time 
when the value becomes invalid and \texttt{Value} the value that is being
stored, which can be of any type.
\begin{figure}[h]
\centering
\begin{lstlisting}
type StorageValue struct {
  Value interface{}
  Time time.Time
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of the \texttt{StorageValue} type.}
\label{fig:StorageValue}
\end{figure}

\subsubsection{Storage}
\texttt{Storage} is an interface representing thread safe mapping from string keys
to \texttt{StorageValues}. Its definition is shown in Figure \ref{fig:Storage} below.
\begin{figure}[h]
\centering
\begin{lstlisting}
type Storage interface {
  Get(string) (StorageValue, bool)
  Set(string, StorageValue)
  Remove(string)
  CompareAndRemove(string, StorageValue) bool
  Keys() []string
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of the \texttt{Storage} interface.}
\label{fig:Storage}
\end{figure}
The package provides a default implementation of this interface that
stores the data in memory using (with minor modifications) sharded concurrent map implementation
by Or Hiltch\footnote{Code available at \url{https://github.com/orcaman/concurrent-map}}.
If a developer wants to store the values in the 
database instead, they can provide their own implementation of the 
\texttt{Store} interface
and use any component that uses the storage the same way as before.
This makes the components that use the storage highly configurable.

\subsubsection{Removing expired values}
In order to free up space used by expired values, the package implements the
\texttt{StorageCleaner} function which does exactly that. Its type signature
is shown below:
\begin{lstlisting}
func StorageCleaner(storage Storage, shutDown <-chan bool, 
                    sleepTime time.Duration)
\end{lstlisting}
The function will repeatedly remove expired values from the provided
\texttt{storage} object until a message is sent on the \texttt{shutDown}
channel. After going through the whole storage object, the cleaner 
sleeps for time specified by the \texttt{sleepTime} argument.
Note that this can be used with any type of storage as it only accesses
the storage object via the methods specified by the \texttt{Storage} 
interface.

\subsection{Caching}
To support caching of outputs of components, the toolkit provides
component generator \texttt{CacheComponent}, which has the following type
signature:
\begin{lstlisting}
func CacheComponent(cache Storage, worker Component, 
                    expiration time.Duration) Component
\end{lstlisting}

\newpage
The returned component behaves as follows:
\begin{itemize}
  \item When an input \texttt{job} holds a request that the component 
        hasn't seen before, then the \texttt{job} is passed to the 
        provided \texttt{worker}. Afterwards the result field of \texttt{job}
        returned by the \texttt{worker} is stored in the provided 
        \texttt{storage} object.

  \item When the component gets a \texttt{job} with a request that it saw 
        previously, it writes the value stored in the \texttt{storage} object,
        to the result field of that \texttt{job} if the value hasn't expired yet. 
        If the value expired, then the component treats the \texttt{job} as if it
        hasn't seen its request before.
\end{itemize}

For the generated component to function properly, the worker must output
a \texttt{job} for every \texttt{job} that is sent to it. This, however, 
is one of the conditions that any component must satisfy, so it shouldn't 
be an issue.

\subsection{Session Management}
This section introduces the component that supports management of any
type of session.
\subsubsection{States}
\label{sec:state}
To implement the logic of the session the toolkit uses \texttt{State}
interface, whose definition is shown in Figure \ref{fig:State}.
\begin{figure}[h]
\centering
\begin{lstlisting}
type State interface {
    Next(job Job) (State, error)
    Terminal() bool
    Result() interface{}
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of the \texttt{State} interface.}
\label{fig:State}
\end{figure}
The expected behavior of the member functions is as follows:
\begin{itemize}
	\item The \texttt{Next} function returns the next state when provided 
          with a \texttt{job} or an error in case the next state can't be 
          generated.
	\item The \texttt{Terminal} function indicates whether the current 
          state is a terminal state.
	\item The \texttt{Result} function returns the result, that possibly 
          after further processing, should be returned to the client.
\end{itemize}

\subsubsection{Session Manager}
\texttt{SessionManager} is a component generator that returns a component
which performs management of sessions for individual users. It has the following signature:
\begin{lstlisting}
func SessionManager(store Storage, initial State, 
                    seshExp time.Duration) Component
\end{lstlisting}
Its arguments are a \texttt{storage} object to be used, an \texttt{initial} 
state for each session and a session expiration time.
The returned component behaves as a session manager. 

The component stores a mapping from session ids to the current state of the session
in the \texttt{storage} object.
Every new request gets assigned a unique session id. Then a current state
of the session is generated from the initial state using the provided
\texttt{job}. Afterwards a mapping from the generated id to the current state
is stored. The result of the call to the \texttt{Result} function on the
current state is then sent down the pipeline and response for the client 
is generated from it.

When a request with set `Session-Id' header comes in, the component gets the current 
state for the session and computes the next state for it based on the provided 
\texttt{job}.
The mapping is then updated with the new state and the result from the new state
is passed down the pipeline.

When a session reaches a terminal state, the session id is removed from the map
and the final result is output.
The states in the mapping are timestamped. Timestamps are updated every time
the state for a session is updated. When a client tries to access an expired
session a new session for this client is started. If the session expiration
time is 0 then the sessions never expire.

% More hype
Note that the abstraction using states allows the users of the 
package to implement the logic for the session, independently of 
the session management part.

\subsection{Dynamic Load Balancing}
To support load balancing based on the current traffic the toolkit implements
\texttt{DynamicLoadBalancer}. It is a component generator whose signature is
shown below. Its architecture is similar
to that of the \texttt{StaticLoadBalancer}.
\begin{lstlisting}
func DynamicLoadBalancer(addTimeout, removeTimeout time.Duration, 
        worker Component, maxWorkers int) Component
\end{lstlisting}
The returned component has an array of workers which can be easily shut down.
The initial number of workers is 1. 
For every input \texttt{job}, the generated component behaves as follows:
\begin{itemize}
	\item It tries to send the \texttt{job} to the workers, if this is not
          successful before the \texttt{addTimeout}, then the component 
          creates a new worker if the current number of workers is smaller 
          than \texttt{maxWorkers}. Then the component tries to send the 
          \texttt{job} to the workers again. This will be successful in 
          case a worker was added because then there is at least one 
          worker that is not busy.

	\item The workers send their results further down the pipeline.

	\item If there are no incoming \texttt{jobs} for duration equal to the 
          \texttt{removeTimeout}, then the component shuts down one 
          of the workers if there is more than one worker.

	\item When the input channel of the component is closed, then 
          the load balancer shuts down all of the workers and afterwards 
          closes its output channel and terminates.
\end{itemize}
There are a lot of other possible load balancing strategies. 
Implementing them should be a simple modification of the existing code.
Note that this is very powerful concept as it can also be used to share
the load between multiple servers and turn them on or shut them down as
required (here using a different strategy might be preferred). There is
also a similar dynamic load balancer for writers.

\subsection{Error handling}

\subsubsection{Error Passer}
A lot of components expect an input of a certain type. Their execution
might end with an error if they are provided with an error result from
the previous component. This would make debugging very hard. So, the 
package provides the following component wrapper, that allows error messages
to bypass a component. The signature of this wrapper is shown below.
\begin{lstlisting}
func ErrorPasser(worker Component) Component
\end{lstlisting}
The inner design of the generated component is shown in Figure 
\ref{fig:errPasserDiag}. The behavior of the generated component 
can be described as follows:
\begin{itemize}
	\item For every input \texttt{job} it checks whether its result field 
          contains an error and if that is the cases it just outputs the 
          \texttt{job}.
	\item If the result filed of the input \texttt{job} doesn't contain 
          an error, then the component passes the \texttt{job} to the worker, 
          which after processing it passes it further down the pipeline.
\end{itemize}

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance = 2cm, auto]
    % Nodes
    \node[component, fill=gray!12, text width=1.5em] (passer) {};
    \node[component, right of=passer] (worker) {Worker};
    \node[text width=2em, right of=worker] (right) {};
    \node[below of=worker, node distance=1.2cm] (lineh) {};
    \node[left of=passer] (left) {};

    % Edges
    \path[line] (left) edge node {in} (passer); 
    \path (right.west) edge node {out} (right.east);
    \path[line] (right.west) -- (right.east);
    \path[line] (passer) -- (worker);
    \path[draw] (worker) -- (right);
    \path[draw] (passer.south) |- (lineh.east);
    \path[draw] (lineh.east) -| (right.west);
    \node[draw, densely dotted, inner sep=1em,
          fit=(passer) (worker) (lineh)] (epasser) {};
    \node[above of=epasser, node distance=1.75cm](name){ErrorPasser};
\end{tikzpicture}
\caption[scale=1.0]{Inner design of the component generated with
  \texttt{ErrorPasser}.}
\label{fig:errPasserDiag}
\end{figure}

\subsubsection{Panic handling}
If we use a component which might cause panic (that is if it can 
throw an exception), this could crash the whole server. 
Hence, the package provides the 
\texttt{PanicHandler} component generator, whose signature is shown 
below.
\begin{lstlisting}
func PanicHandler(worker Component) Component
\end{lstlisting}
It takes a component that can cause panic and returns a component that 
behaves as follows:
\begin{itemize}
	\item It passes every input \texttt{job} to the worker, gets the processed 
        \texttt{job} from it and then passes this \texttt{job} further down
        the pipeline.

	\item In case the worker crashes, that is it causes panic, the component 
        writes an error as a result for the \texttt{job} that caused the crash 
        and then restarts the worker.

	\item The component can itself cause a panic if the worker closes its 
        input channel, or if it closes its output channel before its input 
        channel is closed. That is if the worker violates the contract
        for components.
\end{itemize}

\newpage
\subsection{Network Components}
In this section I present a proxy component, which is an essential
part of a proxy server. Then I introduce three components that
are used to construct the more complex proxy component. These components
are quite useful on their own, especially the network component,
which can be used to delegate tasks to other servers. The other 
two perform rewriting of the http requests and reading of the http
responses.

\subsubsection{Proxy Component}
In order to implement a proxy server one might want to use
a component that forwards the incoming requests to another server
and then reads and outputs the responses. The operation that this
component needs to perform can be divided into three sub-operations: 
rewriting of the request, doing the the request and reading the 
response. To perform each of these sub-operations I implemented
three separate components. These are \texttt{RequestRewriter}, 
\texttt{NetworkComponent} and \texttt{ResponseReader}.

Now the proxy component is just a linear combination of these three
sub-components. Its implementation using the \texttt{LinkComponents}
function is shown in Figure \ref{fig:ProxyComp} below. 
The sub-components are described in the
following sections. Note that all sub-components are wrapped
in an \texttt{ErrorPasser} to avoid doing unnecessary work 
and to make the component more robust.

\begin{figure}[h]
\centering
\begin{lstlisting}
func ProxyComponent(scheme, host string, 
                    client *http.Client) Component {
  return LinkComponents(
    ErrorPasser(RequestCopier(scheme, host)),
    ErrorPasser(NetworkComponent(client)),
    ErrorPasser(ResponseProcessor))
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of Proxy Component generator.}
\label{fig:ProxyComp}
\end{figure}

Note that this approach to building the \texttt{ProxyComponent} exploits
the benefits of the compositional approach. That is we create simple
components each performing a single operation. These can be easily
reused in different scenarios. We can replace parts of the pipeline 
to meet our requirements. If, for example, we wanted to produce different
types of requests, we could replace the \texttt{RequestRewriter} with
a custom component. In case we were only interested in the headers of
the response we could replace the \texttt{ResponseReader} with another 
custom component and so on.

\subsubsection{Request Rewriter}
Request Rewriter is the first part in the pipeline of the 
\texttt{ProxyComponent}. It is implemented as a component generator with the 
following signature:
\begin{lstlisting}
func RequestRewriter(scheme, host string) Component
\end{lstlisting}
It takes string parameters scheme (e.g. http) and host (e.g. www.google.com) 
and returns a component that performs the rewriting. This component
updates the request made by the client that is stored in every input 
\texttt{job} with the given scheme and path. It then stores the 
request in the result field of the given \texttt{job} and outputs it.

\subsubsection{Network Component}
Network Component is the second part in the pipeline of the 
\texttt{ProxyComponent}. It is implemented as a component generator 
with the following signature:
\begin{lstlisting}
func NetworkComponent(client *http.Client) Component
\end{lstlisting}
It takes a single parameter, which is an http client.
The generated component for each input \texttt{job} makes an http request 
taken from the result field of the \texttt{job} using the 
provided http client.
If there is no request in the result field or the request fails, then
the component writes an appropriate \texttt{error} to the result field of the
\texttt{job} and outputs it. If the request is successful then 
the component writes the returned \texttt{http Response} to the 
result field of the \texttt{job} before outputting it.

The behavior of the generated component is described in 
Figure \ref{fig:networkComp} below. This component is usually used
to delegate tasks to other servers. In the case of a proxy server
it delegates all of the work to the other server, but in other scenarios
it can delegate only some tasks or their parts.
\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance = 2cm, auto]
    % Nodes
    \node [component, densely dotted, ellipse] (internet) {Network};
    \node [component, text width=6em, below of=internet, node distance=3cm] (network) {Network Component};
    \node [left of=network, node distance = 3cm] (in) {};
    \node [right of=network, node distance = 3cm] (out) {};
    % Edges
    \path[line] (network) edge [bend right] (internet);
    \path[line] (internet) edge [bend right] (network);
    \path[line] (in) edge node {in} (network);
    \path[line] (network) edge node {out} (out);
\end{tikzpicture}
\caption[scale=1.0]{Design of the network component.}
\label{fig:networkComp}
\end{figure}

Note that the result of the request is fetched in the goroutine where the
component runs. Hence, slow requests might block the pipeline. To avoid
this behavior \texttt{NetworkComponent} is usually used together with one
of the load managers. 

\newpage
\subsubsection{Response Reader}
The \texttt{Response} type of the \texttt{mpserver} package is defined as
a struct that is shown in Figure \ref{fig:Response}.
\begin{figure}[h]
\centering
\begin{lstlisting}
type Response struct {
    Header http.Header
    ResponseCode int
    Body []byte
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of type Response.}
\label{fig:Response}
\end{figure}

It holds the headers, repose code and body of an http response. 
The difference between
this definition and the definition in the default \texttt{http} package is that
the body of my \texttt{Response} is a slice of bytes. 
That is the body has already been
read and the connection to the server closed. This is useful
in the case of a proxy server, as we can cache these objects and reuse
them. This is not possible with the default response type,
as its body can be only read once.

The \texttt{ResponseReader} is a component that for each input \texttt{job}
looks if its result field contains the default \texttt{http} Response. 
If that is the case,
then it reads the response and transfers it to \texttt{mpsever Response} 
type, which it
then writes to the result field and outputs the \texttt{job}. If no response
is provided, then the component writes an appropriate error to the
\texttt{job} before outputting it.

This is the last part in the pipeline of the \texttt{ProxyComponent}.
To use this component after the \texttt{NetworkComponent}, 
the \texttt{ResponseReader} should be wrapped in an error passing 
component, so that when the \texttt{Network\-Component} fails, then 
the \texttt{ResponseReader} won't have to do any work.


\subsection{Summary}
In this chapter I introduced more advanced component wrappers and generators
and described how to use them. The next chapter shows a few example
servers implemented using my toolkit.

