\section{Implementation}
\label{sec:impl}
In this Chapter I present basic elements of the \texttt{mpserver} toolkit
and how to work with them. The most important parts are the primitive types:
\texttt{Job}, \texttt{Component} and \texttt{Writer}.

\subsection{Jobs}
\texttt{Job} is an interface that represents the type of messages that are
passed between the components of the network. Its definition is shown
in Figure \ref{fig:Job} below.

\begin{figure}[h]
\centering
\begin{lstlisting}
type Job interface {
    //Public methods
    GetRequest() *http.Request
    GetResult() interface{}
    SetResult(value interface{})
    SetResponseCode(respCode int)
    SetResponseCodeIfUndef(respCode int)
    SetHeader(key string, value string)

    // Private methods
    getResponseWriter() http.ResponseWriter
    getResponseCode() int
    writeHeader()
    write(body []byte)
    close()
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of the type \texttt{Job}.}
\label{fig:Job}
\end{figure}

Note that in Go only types, functions, constants and variables whose name
starts with a capital letter are exported from a package. Hence, anyone
using the package can only use the first 6 methods of the \texttt{Job}
interface.

The interface is implemented by a pointer to a private struct type shown in 
Figure \ref{fig:jobStruct}.
\newpage
\begin{figure}[h]
\centering
\begin{lstlisting}
type jobStruct struct {
    request *http.Request
    result interface{}

    responseCode int
    responseWriter http.ResponseWriter
    done chan<- bool
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of a private type that implements the 
\texttt{Job} interface.}
\label{fig:jobStruct}
\end{figure}
Here:
\begin{itemize}
  \item \texttt{request} is a pointer to the HTTP request made by the 
        client implemented in the default http library. It can be accessed
        using the \texttt{GetRequest} method. 

  \item \texttt{result} is the result computed so far (initially \texttt{nil}). This is
        of type \texttt{interface\{\}} which is implemented by any type.
        Hence, values of all types can be stored there. This field can be
        accessed and changed using the \texttt{GetResult} and \texttt{SetResult}
        methods respectively.
  
  \item \texttt{responseCode} is the HTTP response code that should be 
        returned to the client. The default value of this field is $-1$,
        which indicates that the \texttt{responseCode} hasn't been set yet. 
        It can be set using the \texttt{SetResponseCode} or 
        \texttt{SetResponseCodeIfUndef} methods, where the second method only
        makes a change in case the response code has not been set yet. The value
        of the response code can be retrieved using the \texttt{getResponseCode}
        method inside the package. This is only used by a 
        \texttt{Writer} type that is introduced later.

  \item \texttt{responseWriter} is an object that writes the response back 
        to the client. Again only \texttt{Writers} use this field. They
        can retrieve it using \texttt{getResponseWriter} method. 
        Alternatively, they can use the \texttt{write\-Header}
        and \texttt{write} methods to access the field indirectly.
        These methods write the headers and 
        the body of the response to the client respectively.

  \item \texttt{done} is a signaling channel. Signal is sent on this channel,
		after the response is written at the end of the pipeline using the 
        \texttt{close} method. This allows the connection to the client to 
        be closed after the result is written.
\end{itemize}
Note that all methods related to writing the response back to the client
are private to the package. This is to ensure correct behavior, as otherwise
these methods could be accessed by users of the package in any part of the 
network, what could lead to panic (exceptions).

\subsection{Components}
This section introduces the \texttt{Component} type and then shows how we can simply
create components. Afterwards it presents a way how multiple components can
be combined to form a linear pipeline. The sections ends with a description
of a component generator that implements simple load balancing among
multiple components.

\subsubsection{The Component type}
Component is a generic part of the pipeline with an input and 
an output channel. It reads \texttt{jobs} from its input channel, 
processes them and writes the updated \texttt{jobs} to the output 
channel. The type \texttt{Component} is defined as follows:
\begin{lstlisting}
type Component func (in <-chan Value, out chan<- Value)
\end{lstlisting}
Every component should satisfy the following conditions:
\begin{itemize}
    \item When run a component should eventually read all \texttt{jobs} that are
          written to its input channel.

    \item Every \texttt{job} that a component reads from its input channel
          should be written to its output channel after the \texttt{job} was processed
          and possibly modified. This ensures that every request gets a response.

    \item A component can terminate only when its input channel is closed. 
          Before terminating, the component should close its output channel.
          This ensures termination of any components further down the network.

    \item Component can only close its output channel and it can only do so
    	  after its input channel have been closed. It should never attempt
          to close the input channel as this will likely result in a panic.
\end{itemize}

\subsubsection{Making Components}
To simplify the construction of components, so that they adhere to 
the above specification the package provides a helper function that 
constructs a component using a function that takes a \texttt{Job} object 
and processes it. The object can be modified by the function as it is 
implemented by a pointer type. The function has the following signature:
\begin{lstlisting}
type ComponentFunc func (job Job)
\end{lstlisting}
This represents a transition function that is applied to input \texttt{jobs}
to produce output \texttt{jobs}.
The implementation of the described component generator is shown in 
Figure \ref{fig:MakeComponent} below.
\begin{figure}[h]
\centering
\begin{lstlisting}
func MakeComponent(f ComponentFunc) Component {
    return func (in <-chan Job, out chan<- Job) {
        for job := range in {
            f(job); out <- job
        }
        close(out)
    }
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of the \texttt{MakeComponent} function.}
\label{fig:MakeComponent}
\end{figure}

Here the for loop reads \texttt{jobs} from the input channel and assigns them
to the variable \texttt{job} while the input channel is open. Then the 
the function \texttt{f} is applied to this \texttt{job} and finally 
the \texttt{job} is outputted.
Afterwards, the component waits for another \texttt{job}. When the input channel
is closed the loop terminates. Afterwards, the component closes its output channel
and terminates.

\subsubsection{Constant Component}
Constant Component is a simple component generator that takes a value \texttt{c}
of any type and returns a component that writes \texttt{c} to the result field of 
all input \texttt{jobs} and then outputs them. It's definition using the 
\texttt{MakeComponent} generator is shown in Figure 
\ref{fig:ConstantComponent} below.
\begin{figure}[h]
\centering
\begin{lstlisting}
func ConstantComponent(c interface{}) Component {
    return MakeComponent(func (job Job) {
        job.SetResult(c)
    })
}
\end{lstlisting}
\caption[scale=1.0]{Declaration of constant component generator.}
\label{fig:ConstantComponent}
\end{figure}

\subsubsection{Linking components}
The library also provides a component generator \texttt{LinkComponents} that takes 
any number of components and returns a component that behaves as their 
linear combination. That is as a pipeline constructed from these components 
in the order in which they are provided. Its signature is shown below.
\begin{lstlisting}
func LinkComponents(components ...Component) Component
\end{lstlisting}
The subcomponents are linked and run when the combined component is run. 
The last of the provided components is run in the goroutine in which
the combined component was run.

\subsubsection{Simple Load Balancing}
To support running the same component with the same input and output 
channels multiple times the toolkit provides \texttt{StaticLoadBalancer}.
Note that this is possible as channels in go are many-to-many.
\texttt{StaticLoadBalancer} is a component generator that takes a worker 
component and number $n$ and returns a component that when run starts 
$n$ instances of the worker component. 

The case with $n=3$ in a simple pipeline is shown in Figure \ref{fig:slbEx} below. 
In order to support clean shutdown of the combined component, the load balancer uses 
wrappers (U shaped components in the diagram) around workers and a component that 
forwards incoming \texttt{jobs} to the workers. When input channel of the forwarding
component is closed it sends a signal to the component wrappers on shut 
down channels (dashed arrows in the diagram). The wrappers then close
the channels to the workers, which will force the workers to terminate.
Finally the distributor closes the output channel of the whole component
and then the termination signal propagates further down the network.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance = 1.5cm, auto,
    workerWidth/.style={text width=5em, text centered},
    workerHeight/.style={minimum height=4em},
    UWidth/.style={minimum width=0.6cm},
    UHeight/.style={minimum height=0.6cm}
    ]
    % Nodes
    \node [component] (worker1) {Worker};
    \node [workerWidth, UHeight, below of=worker1] (passer1) {};
    \node [workerWidth, UHeight, above of=worker1] (passer2) {};
    \node [component, above of=passer2] (worker2) {Worker};
    \node [component, below of=passer1] (worker3) {Worker};
    \node [workerWidth, UHeight, below of=worker3] (passer3) {};
    \node [above of=worker2, node distance=1.3cm, xshift=-0.8cm] {Simple Load Balancer};

    % Left
    \node [UWidth, workerHeight, left of=worker1, node distance=2cm] (left1) {};
    \node [UWidth, workerHeight, left of=worker2, node distance=2cm] (left2) {};
    \node [UWidth, workerHeight, left of=worker3, node distance=2cm] (left3) {};

    \node [UWidth, workerHeight, left of=left1, node distance=1.5cm] (lleft1) {};
    \node [UWidth, workerHeight, left of=left2, node distance=1.5cm] (lleft2) {};
    \node [UWidth, workerHeight, left of=left3, node distance=1.5cm] (lleft3) {};

    \draw[fill=gray!12] (lleft2.north west) -- (lleft2.west |- passer3.south) --
          (lleft2.east |- passer3.south) -- (lleft2.north east) --
          cycle;

    % Right
    \node [UWidth, workerHeight, right of=worker1, node distance=2cm] (right1) {};
    \node [UWidth, workerHeight, right of=worker2, node distance=2cm] (right2) {};
    \node [UWidth, workerHeight, right of=worker3, node distance=2cm] (right3) {};

    \draw[fill=gray!12] (left1.north west) -- (left1.west |- passer1.south) -- 
          (right1.east |- passer1.south) -- (right1.north east) -- 
          (right1.north west) -- (right1.west |- passer1.north) -- 
          (left1.east |- passer1.north) -- (left1.north east) -- cycle;
    \draw[fill=gray!12] (left2.north west) -- (left2.west |- passer2.south) -- 
          (right2.east |- passer2.south) -- (right2.north east) -- 
          (right2.north west) -- (right2.west |- passer2.north) -- 
          (left2.east |- passer2.north) -- (left2.north east) -- cycle;
    \draw[fill=gray!12] (left3.north west) -- (left3.west |- passer3.south) -- 
          (right3.east |- passer3.south) -- (right3.north east) -- 
          (right3.north west) -- (right3.west |- passer3.north) -- 
          (left3.east |- passer3.north) -- (left3.north east) -- cycle;

    \node [listener, left of=left1, node distance=4cm] (listener) {Listener};
    \node [writer, right of=right1, node distance=3cm] (writer) {Writer};
    % Edges
    \path [line] (listener) -- (lleft1);
    \path [line] (lleft1) -- (left1);
    \path [line] ($ (lleft1) !.4! (left1) $) |- (left2);
    \path [line] ($ (lleft1) !.4! (left1) $) |- (left3);

    \path [line, dashed,transform canvas={shift={(0,-0.5)}}] (lleft1) -- (left1);
    \path [line, dashed,transform canvas={shift={(0,-0.5)}}] (lleft2) -- (left2);
    \path [line, dashed,transform canvas={shift={(0,-0.5)}}] (lleft3) -- (left3);

    \path [line] (left1) -- (worker1);
    \path [line] (left2) -- (worker2);
    \path [line] (left3) -- (worker3);

    \path [line] (worker1) -- (right1);
    \path [line] (worker2) -- (right2);
    \path [line] (worker3) -- (right3);
    \path [line] (right1) -- (writer);
    \path [draw] (right2) -| ($ (right1) !.3! (writer) $);
    \path [draw] (right3) -| ($ (right1) !.3! (writer) $);
    \node[xshift=0.5em, draw, densely dotted, minimum width=18em, inner sep=0.5em,fit=(right2) (right3) (lleft2) (lleft3) (passer3)] {};
\end{tikzpicture}
\caption[scale=1.0]{Simple server example showing usage of \texttt{StaticLoadBalancer}.}
\label{fig:slbEx}
\end{figure}

This combined component allows multiple requests to be processed by the worker
components at the same time and this in turn allows faster requests to 
overtake slower requests in the pipeline. The worker component can represent
a simple computation done locally or they can connect to other servers
and perform expensive computation there. More advanced version of load 
balancing component is introduced in the next chapter.

\newpage
\subsection{Writers}
In this section I introduce the \texttt{Writer} type, which acts as the end of the 
pipeline. Then I present a way how to create custom writers.
\subsubsection{Writer type}
Writer is the end of the pipeline which writes results back to the client.
The type \texttt{Writer} is defined as follows:
\begin{lstlisting}
type Writer func (in <-chan Job)
\end{lstlisting}
It is a function with only an input channel.
Every writer satisfies the following specification.
\begin{itemize}
    \item When run a writer eventually reads all \texttt{jobs} that are written 
          to its input channel.

	\item For every \texttt{job} that is  read from the input channel a writer 
          behaves as follows. It firstly checks if the result field 
          contains the expected type. If that is the case, it first 
          writes the headers and then the body of the response using 
          the \texttt{writeHeader} and \texttt{write} methods of the \texttt{job} respectively. Afterwards it closes the connection for the \texttt{job} 
          using the \texttt{close} method. If the value in the result 
          field has an incorrect type, the writer writes an appropriate 
          error message to the client and closes the connection for the 
          \texttt{job}. 

	\item The writer terminates when its input channel is closed.
\end{itemize}
The package provides writers for writing string, json, file, error and many
other types of responses. There is also a writer generator for static
load balancing similar to the one for components. It's implementation
is significantly simpler as we don't have to forward the processed
\texttt{jobs} further down the network. Note that components
combined with writers appear like a single writer. Hence, we can use the
load balancers for writers to manage a more complex part the network.

\subsubsection{Making Writers}
As noted before the methods for writing responses from \texttt{jobs}
back to the client cannot be used outside of the package. To allow
for construction of custom writers the package provides writer generator
component similar to the \texttt{MakeComponent} function. It uses 
functions with the following signature to generate the writers.
\begin{lstlisting}
type WriterFunc func (job Job) ([]byte, error)
\end{lstlisting}
This function should set headers and convert the provided \texttt{job} into 
a slice\footnote{``Slice is a dynamically-sized, flexible view into the elements 
of an array.''\cite{tour}} of bytes or return an error in case it 
is provided with a 
wrong input. The implementation of the writer generator itself is shown
in Figure \ref{fig:MakeWriter} below.
\begin{figure}[h]
\centering
\begin{lstlisting}
func MakeWriter(writerFunc WriterFunc) Writer {
    return func (in <-chan Job) {
        for job := range in {
            resp, err := writerFunc(job)

            if (err == nil) {
                // Write the response
                job.SetResponseCodeIfUndef(http.StatusOK)
                job.writeHeader()
                job.write(resp)
                job.close()
            } else {
                writeError(job, err)
            }
        }       
    }
}
\end{lstlisting}
\caption[scale=1.0]{Implementation of writer generator.}
\label{fig:MakeWriter}
\end{figure}

The generated writer runs the provided \texttt{writerFunc} for every \texttt{job} 
that is inputted. If the function returns an error it writes the error 
back to the client using the \texttt{writeError}\footnote{See Appendix 
\ref{sec:writer.go} for the implementation.} function. Otherwise, it 
writes the headers and the body of the response to the client and then 
closes the connection. (The \texttt{writeError} does this for the error
results).

\subsection{Router and Collector}
In this section I introduce a way to split a stream of incoming \texttt{jobs} into
multiple streams of \texttt{jobs} that go into different parts of the network
using very flexible rules for routing. Then I present a way how to
safely combine multiple streams into one.

\subsubsection{Conditions}
Condition type is defined as follows:
\begin{lstlisting}
type Condition func (job Job) bool
\end{lstlisting}
It is a just function that takes a \texttt{job} and returns a boolean.
The function shouldn't change the provided \texttt{jobs}.

\subsubsection{Router}
Conditions are used in Router, which is a component with the 
following signature:
\begin{lstlisting}
func Router(in <-chan Job, defOut chan<- Job, 
            outs []chan<- Job, conds []Condition)
\end{lstlisting}
It takes an input channel, default output channel, a slice of output channels 
and a slice of conditions.
The number of output channels and the number of conditions should be the same.
For every input \texttt{job} the Router evaluates the conditions from first to last.
When a condition returns \texttt{true} the \texttt{job} is written to a 
corresponding output channel 
and the processing of the current \texttt{job} terminates. If all conditions 
return \texttt{false} then the \texttt{job} is written to the default output channel.

\subsubsection{Error Router}
The package also implements a special router that redirects error messages
to a special error channel and sends all other messages to the default output 
channel. This is used to avoid sending \texttt{jobs} that hold errors to 
any other type of writer than the error writer. If we allowed this, then
all response errors will only show that a particular writer got a wrong input.
The signature of the \texttt{ErrorRouter} function is shown below.
\begin{lstlisting}
func ErrorRouter(in <-chan Job, defOut chan<- Job, 
                                errChan chan<- Job)
\end{lstlisting}

\subsubsection{Collector}
Collector is a component whose behavior is the opposite of the behavior 
of a router. It gets multiple input channels
and forwards their outputs to a single output channel. Now as channels
in go are many-to-many we could just use a single channel to which
multiple components would output the \texttt{jobs}. 

However, when we shutdown
one of the components, this component will close the output channel
what will cause panic next time one of the other components tries to write
to it. Now, \texttt{Collector} only closes its output channel, when
all of its input channels have been closed, thus avoiding the problem 
described above. It is acceptable to just use a single channel instead of
the collector component, if we don't intend to shut down the network.
The type signature of the \texttt{Collector} is show below.
\begin{lstlisting}
func Collector(ins []<-chan Job, out chan<- Job)
\end{lstlisting}


\subsection{Summary}
In this Chapter I introduced the basic elements of the toolkit. This 
is enough to implement some simple serves. The next section introduces
more advanced components that allow more complex behaviors.

